%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% nuthesis-template.tex - Miguel A, Lerma - 5/21/2012
%                         mlerma@math.northwestern.edu
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%% DISCLAIMER %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
% In spite of the effort to accommodate this template and the nuthesis
% class to the official requirements of the university to write a 
% Ph.D. dissertation, it is not possible to guarantee that it will 
% always work, and the author of the dissertation remains responsible
% for checking that such requirements are actually fulfilled by 
% his/her final work.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\documentclass[12pt]{nuthesis}	% The nuthesis class is based on 
				% amsbook.cls.			
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{caption}
\usepackage{epigraph}
\usepackage{subcaption}
\usepackage{amssymb,amsmath,amsbsy}
\usepackage{multirow}
\usepackage{subeqnarray}
\usepackage{color}
\definecolor{codegreen}{rgb}{0,0.6,0}
\usepackage{listings}
\definecolor{listinggray}{gray}{0.9}
\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
\lstset{
  backgroundcolor=\color{lbcolor},
  tabsize=2,
  language=C++,
  captionpos=b,
  tabsize=3,
  frame=lines,
  numbers=left,
  numberstyle=\tiny,
  numbersep=5pt,
  breaklines=true,
  showstringspaces=false,
  basicstyle=\small\sf,
%  identifierstyle=\color{magenta},
  keywordstyle=\color[rgb]{0,0,1},
  commentstyle=\color{codegreen},
  stringstyle=\color{red}
}
%\usepackage[style=authoryear,backend=bibtex]{biblatex}
%\bibliography{references}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DATA OF AUTHOR AND DISSERTATION %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Mao Mao}

\title{Electroosmotic flow in a nanopore}

\degree{DOCTOR OF PHILOSOPHY}  % Default: DOCTOR OF PHILOSOPHY

\field{Mechanical Engineering}            % Default: Mathematics

\graduationmonth{March}         % The default is June or December
                                % depending on current date.
\graduationyear{2015}          % Default: current year.

\begin{document}
%	
%	THE BODY OF YOUR THESIS STARTS HERE
%

%%%%%%%%%%%%%%%%%%%%%%
% Some initial stuff %
%%%%%%%%%%%%%%%%%%%%%%

\frontmatter		% Preliminary pages start here.

\maketitle		% Produces the title page.

\copyrightpage		% Creates the copyright page.

\abstract		% Abstract.
Nanopores have very important applications in molecule\//polymer detection and separation. The flow field in the vicinity of a nanopore is crucial to the motion of the translocating molecule\//polymer. This dissertation addresses the problem of electroosmotic pumping of fluid through a nanopore that traverses a charged, dielectric membrane. 

The strength and structure of the flow depends on several parameters including the geometry of the nanopore, external voltage, concentration of the electrolyte solution and the properties of the membrane material, such as polarity and surface charge density. Under weak field assumption, perturbation method can be utilized for theoretical modeling of flow field in the linear regime. When the membrane thickness is zero, the reciprocal theorem gives the flow rate generated by an applied voltage in the form of an integral over the fluid domain. For a circular hole in a membrane of zero thickness, the integral can be worked out in closed form and computed up to quadrature. When the membrane thickness is non--zero, the flow rate under applied field can be obtained by patching the zero--thickness solution with the analytical solution of electroosmotic flow in a cylindrical channel. The effect of the electric double layer thickness is also considered and theory is modified when it is not negligible compared to the membrane thickness. When the intrinsic surface charges on the membrane are neutralized, eddies with length scale of a few nanometers could arise within the pore due to induced charge electroosmosis. The flow strength in this case is found to be proportional to square of the applied voltage. The shape of the eddies is found to be dependent on the geometry of the pore.

Numerical solvers to solve the full governing system of equations are developed based on finite volume method. The solvers are able to compute the coupled electrostatics, ionic transport and Stokes flow equations. Numerical results agree with theoretical models in the corresponding parameter regimes. A finite element software package is also utilized to capture the multi--physics processes. The simulation tools are applied to study electroosmotic nanojet from a nanocapillary, and the results are compared with the experimental work at our collaborator\'{}s group.

\acknowledgements	% Acknowledgements (optional).
I would like to express my deepest gratitude to my advisor Prof. Sandip Ghosal. He has been a tremendous mentor for me during my years at Northwestern University. I have learned from him a great deal, not only about fluid mechanics, but also about commitment to scientific research. Working with him has always been a pleasant and rewarding experience. His support and advice on both research as well as my career has been invaluable. So with all my heart, thank you Professor, for everything!

I would also like to thank Dr. John Sherwood from DAMTP at the University of Cambridge. The material presented in this dissertation would be impossible without his extremely important contributions. It was a great honor collaborating with a great mathematician like him. I also want to thank Dr. Guohui Hu from Shanghai University, who introduced the topic of nanopores to me in 2011. The research we did together on nanopores at the beginning of my PhD study pointed me in the right path.

I would like to thank my committee members, Prof. Seth Lichter, Prof. Malcolm MacIver and Prof. Michael Miksis, for taking the time to review and comment on my thesis. I also want to thank Prof. Neelesh Patankar for helpful discussion on computational fluid dynamics, and Prof. Ulrich Keyser for hospitality during my visit to the Cavendish Laboratory at Cambridge.

Most importantly, I would like to thank my dearest girl friend Anna, for being with me through this tough journey, for sharing all my happy and sad moments, and for loving and unconditionally supporting me during the past 3 years. I am also grateful to my family who encouraged and support me in pursuing my PhD studies.

A special thanks to the guys in the fluids lab, Amneet, Namu, Paul, Rahul, Tayil, Tom, Walter and Zhen, and my good friends Hao, Hongyi, Jieqi, Shi, Shu and Wenchang.

I have been extremely fortunate to have all the help and support I have had during the past 5 years. I am grateful to everyone who helped me along the road, and I wish I were able to name each one of them but this would take too long as there were many! 

%\preface		% Preface (optional).
%
%This is the preface.

%% A few more optional pages (uncomment if needed)
%
%\listofabbreviations 
%
%This is the list of abbreviations (optional).
%
%\glossary
%
%This is the glossary (optional).
%
%\nomenclature
%
%This is the nomenclature (optional).
%
%% Note that the dedication text must be passed as an argument
%% of the \dedication command
%\dedication{This is the dedication (optional).}
%

\clearpage\phantomsection % needed for the hyperlinks to work correctly
\tableofcontents	% Table of Contents will be automatically
			% generated and placed here.

\clearpage\phantomsection % needed for the hyperlinks to work correctly
\listoftables		% List of Tables and List of Figures will be placed

\clearpage\phantomsection % needed for the hyperlinks to work correctly
\listoffigures		% here, if applicable (optional).



\mainmatter             % Actual text starts here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Actual text starts here %
%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Alternatively, you may write the chapters in separate
% files, say chap1.tex, chap2.tex, etc., and include them 
% with commands:

\include{introduction/introduction}
\include{zero_thickness/zero_thickness} 
\include{finite_thickness/finite_thickness}
\include{eddies/eddy8}
\include{openfoam/openfoam}
\include{conclusion/conclusion}

% The command \includeonly above allows to include a selected 
% set of chapters only.

% (The following suggested by Francisco Iacobelli - 5/11/2010)
% In case, you want to use BibTeX, you should replace (or comment)
% the bibliography environment.
% Instead uncomment the following 3 lines and replace <bib file>
% with your .bib file:
%\renewcommand\refname{\begin{centering}References\end{centering}}
\bibliography{references}
\bibliographystyle{unsrt} %or another suitable style.

%\appendix		% Appendix begins here (optional).
%
%%\appendices	        % If more than one appendix chapters,
%				% use appendices instead of appendix
%
%\chapter{Title of Appendix}	% First appendix chapter, i.e., Appendix A.
%
%\section{First section of Appendix}	% This is appendix section 1.
%
%This is the Appendix.
%
\end{document}

